---
title: "Clinica full QC metrics visualization"
author: "Ju-Chi.Yu"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE, message='hide', warning='hide'}
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(readr))
suppressMessages(library(ExPosition))
suppressMessages(library(PTCA4CATA))
suppressMessages(library(data4PCCAR))
suppressMessages(library(dendextend))
```

## Read data

```{r, include = FALSE}
data.path <- "../data/"
oasis_dice <- read_tsv(paste0(data.path, "oasis_metrics.tsv")) %>% as.data.frame
oasis_dice_num <- oasis_dice[,c(4,5,7:12)]

```

```{r}
head(oasis_dice)
```

## Scatter plot {.tabset}

The scatter plots illustrate the relationship between each metric and DARQ probability.

### dice BET

The correlation between the two metrics is `r cor(oasis_dice$dice_bet_probability, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = dice_bet_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### dice HD-BET

The correlation between the two metrics is `r cor(oasis_dice$dice_hd_bet_probability, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = dice_hd_bet_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Mutual information

The correlation between the two metrics is `r cor(oasis_dice$mutual_info, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = mutual_info, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Correlation ratio

The correlation between the two metrics is `r cor(oasis_dice$correlation_ratio, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = correlation_ratio, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Norm mi

The correlation between the two metrics is `r cor(oasis_dice$norm_mi, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = norm_mi, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Correlation coefficient

The correlation between the two metrics is `r cor(oasis_dice$correlation_coef, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = correlation_coef, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### L1-norm correlation coefficient

The correlation between the two metrics is `r cor(oasis_dice$cr_l1, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = cr_l1, y = darq_probability, color = darq_pass)) +
  geom_point()
```

## Correlation plot

This figure shows how these methods are correlated to one another. In general, everything is positively correlated meaning that they are capturing similar information. The darker squares, however, illustrates the metrics that cluster together based on their similarity.

```{r}
corrplot::corrplot(cor(oasis_dice_num), method = "shade", order = "hclust", hclust.method = "ward.D2")
```

## Explore

To explore the relationship and compare between the metrics, we first look at their distributions, then perform a principal component analysis (PCA) on all metrics.

```{r, echo = FALSE}
color.ind <- c("TRUE" = "#228B22",
               "FALSE" = "#B22222")
color.ind.all <- recode(as.character(oasis_dice$darq_pass), !!!color.ind)
```

### Density plot

The distribution showed that DARQ has very different distribution as compared to other metrics as it tends to give extreme scores. 

```{r}
oasis_dice_num %>% pivot_longer(colnames(oasis_dice_num)) %>%
  ggplot(aes(x = value)) + geom_histogram(bins = 25) + facet_wrap(~name)
```

### PCA

We performed a PCA on all metrics aside from DARQ to examine the relationships between them. Because these metrics are all between 0 and 1, we performed to run a non-centered-non-scaled PCA to keep the variance and the values of the metrics. This PCA is equivalent to performing a singular value decomposition directly on the data.

```{r, echo = FALSE}
## run PCA
PCAres <- epPCA(oasis_dice_num[,-2], 
                # DESIGN = oasis_dice$darq_pass,
                center = FALSE,
                scale = FALSE,
                graphs = FALSE)

map.label <- createxyLabels.gen(lambda = PCAres$ExPosition.Data$eigs,
                                tau = PCAres$ExPosition.Data$t)
```

```{r, echo = FALSE}
## Plot factor scores
fi.map <- createFactorMap(PCAres$ExPosition.Data$fi,
                          col.points = color.ind.all,
                          alpha.points = 0.3,
                          col.background = NULL,
                          col.axes = "black")

fi.map$zeMap_background + fi.map$zeMap_dots + map.label
```

The row factor scores (participants) showed a separation roughly matching the pass/fail outcomes from DARQ. In the figure, GREEN dots indicate passing participants, and RED dots indicate failing ones.

```{r, echo = FALSE}
## Column factor scores
fj.map <- createFactorMap(PCAres$ExPosition.Data$fj,
                          col.background = NULL,
                          col.axes = "black")

fj.arrow <- addArrows(PCAres$ExPosition.Data$fj)

fj.map$zeMap + fj.arrow + map.label + coord_cartesian()
```

From the column factor scores (metrics), we identified three clusters of metrics. To narrow it down to few metrics, we chose `dice_hd_bet_probability` which we think is the most meaningful to include, and `correlation_coef` which is less related to the other metrics (as indicated by large angle between all metrics and itself).

### Grouping with the two metrics

From `dice_hd_bet_probability` and `correlation_coef`, we examined their relationships and try to derive different QC groups.

#### Distributions

We first checked the distribution of the two metrics and used these distribution to identify a cut-off point. For DICE_HD_BET, we set it to 0.89, and for correlation coefficient, we set it to 0.5 (indicated by the red lines).

```{r}
FinMetric <- oasis_dice %>% select(dice_hd_bet_probability, correlation_coef) %>% as.data.frame
rownames(FinMetric) <- oasis_dice$participant_id

FinMetric %>% ggplot(aes(dice_hd_bet_probability)) + geom_histogram(bins = 30) + geom_vline(xintercept = 0.89, color = "red", lwd = 2)
FinMetric %>% ggplot(aes(correlation_coef)) + geom_histogram(bins = 25) + geom_vline(xintercept = 0.5, color = "red", lwd = 2)

```

#### scatter plot

We plotted the scatter plot between the two current metrics with their newly-derived pass/fail outcomes.

The correlation between the two metrics is `r cor(FinMetric$correlation_coef, FinMetric$dice_hd_bet_probability)`

-   The group labels are in (`Pass/Fail_corr_coef`)\_(`Pass/Fail_dice_hd_bet`)

```{r, echo = FALSE, message=FALSE, warning=FALSE}
## Pass/Fail based on two metrices
FinMetric$corr_ratio_group <- FinMetric$correlation_coef > 0.5
FinMetric$dice_hdbet_group <- FinMetric$dice_hd_bet_probability > 0.89

## Combine the two metrics
FinMetric$both_group <- sprintf("%s_%s", FinMetric$corr_ratio_group, FinMetric$dice_hdbet_group)

col.both.group <- c("TRUE_TRUE" = "#228B22",
                    "TRUE_FALSE" = "#20B2AA",
                    "FALSE_TRUE" = "#6959CD",
                    "FALSE_FALSE" = "#B22222")
FinMetric$both_group_col <- recode(FinMetric$both_group, !!!col.both.group)

library(plotly)
plot_ly(data = FinMetric, 
        x = ~correlation_coef, 
        y = ~dice_hd_bet_probability, 
        color = ~both_group, 
        colors = col.both.group,
        text = ~rownames(FinMetric))

# ggplot(FinMetric, 
#        aes(x = correlation_coef, y = dice_hd_bet_probability, color = both_group)) +
#   geom_point() +
#   scale_color_manual(values = col.both.group)

```

#### Participants in 4 groups {.tabset}

##### Pass Both

```{r}
FinMetric %>% filter(both_group == "TRUE_TRUE") %>% rownames
```

##### Fail Both

```{r}
FinMetric %>% filter(both_group == "FALSE_FALSE") %>% rownames
```

##### Pass Corr_coef Fail dice_hd_bet

```{r}
FinMetric %>% filter(both_group == "TRUE_FALSE") %>% rownames
```

##### Fail Corr_coef Pass dice_hd_bet

```{r}
FinMetric %>% filter(both_group == "FALSE_TRUE") %>% rownames
```
