---
title: "Clinica full QC metrics visualization"
author: "Ju-Chi.Yu"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE, message='hide', warning='hide'}
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(readr))
suppressMessages(library(ExPosition))
suppressMessages(library(PTCA4CATA))
suppressMessages(library(data4PCCAR))
suppressMessages(library(dendextend))
```

## Read data
```{r, include = FALSE}
data.path <- "../data/"
oasis_dice <- read_tsv(paste0(data.path, "oasis_metrics.tsv"))
oasis_dice_num <- oasis_dice[,c(4,5,7:12)]

```

```{r}
head(oasis_dice)
```

## Scatter plot {.tabset}

### dice BET

The correlation between the two metrics is `r cor(oasis_dice$dice_bet_probability, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = dice_bet_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### dice HD-BET

The correlation between the two metrics is `r cor(oasis_dice$dice_hd_bet_probability, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = dice_hd_bet_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```
### Mutual information

The correlation between the two metrics is `r cor(oasis_dice$mutual_info, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = mutual_info, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Correlation ratio

The correlation between the two metrics is `r cor(oasis_dice$correlation_ratio, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = correlation_ratio, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Norm mi

The correlation between the two metrics is `r cor(oasis_dice$norm_mi, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = norm_mi, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### Correlation coefficient

The correlation between the two metrics is `r cor(oasis_dice$correlation_coef, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = correlation_coef, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### L1-norm correlation coefficient

The correlation between the two metrics is `r cor(oasis_dice$cr_l1, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = cr_l1, y = darq_probability, color = darq_pass)) +
  geom_point()
```

## Correlation plot
```{r}
corrplot::corrplot(cor(oasis_dice_num), method = "shade", order = "hclust", hclust.method = "ward.D2")
```


## Explore
What if we only look at dice?

### Clustering results
```{r, echo = FALSE}
# install.packages("dendextend")

# color.ind <- c("TRUE_TRUE" = "#228B22", 
#                "TRUE_FALSE" = "#20B2AA", 
#                "FALSE_TRUE" = "#6959CD", 
#                "FALSE_FALSE" = "#B22222")
color.ind <- c("TRUE" = "#228B22",
               "FALSE" = "#B22222")
color.ind.all <- recode(as.character(oasis_dice$darq_pass), !!!color.ind)
```

### Density plot
```{r}
oasis_dice_num %>% pivot_longer(colnames(oasis_dice_num)) %>%
ggplot(aes(x = value)) + geom_histogram(bins = 25) + facet_wrap(~name)
```


### PCA
```{r, echo = FALSE}
## run PCA
PCAres <- epPCA(oasis_dice_num, 
                # DESIGN = oasis_dice$darq_pass,
                center = FALSE,
                scale = FALSE,
                graphs = FALSE)

map.label <- createxyLabels.gen(lambda = PCAres$ExPosition.Data$eigs,
                                tau = PCAres$ExPosition.Data$t)
```

```{r}
## Plot factor scores
fi.map <- createFactorMap(PCAres$ExPosition.Data$fi,
                          col.points = color.ind.all,
                          alpha.points = 0.3,
                          col.background = NULL,
                          col.axes = "black")

fi.map$zeMap_background + fi.map$zeMap_dots + map.label

## Column factor scores
fj.map <- createFactorMap(PCAres$ExPosition.Data$fj,
                          col.background = NULL,
                          col.axes = "black")

fj.arrow <- addArrows(PCAres$ExPosition.Data$fj)

fj.map$zeMap + fj.arrow + map.label
```

If we narrow it down to a subset, we would choose `dice_hd_bet_probability`, `cr_l1`, `correlation_ratio`, and `norm_mi`


