---
title: "Clinica QC metrics visualization"
author: "Ju-Chi.Yu"
date: "2023-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE, message='hide', warning='hide'}
suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(readr))
suppressMessages(library(ExPosition))
suppressMessages(library(PTCA4CATA))
suppressMessages(library(data4PCCAR))
suppressMessages(library(dendextend))
```

## Read data
```{r, include = FALSE}
data.path <- "../data/"
oasis_dice <- read_tsv(paste0(data.path, "oasis_dice.tsv"))
```

```{r}
head(oasis_dice)
```

## Visualize {.tabset}

### scatter plot - all

The correlation between the two metrics is `r cor(oasis_dice$dice_probability, oasis_dice$darq_probability)`

```{r, echo = FALSE}
ggplot(oasis_dice, 
       aes(x = dice_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```

### scatter plot - failed

```{r, echo=FALSE}
oasis_dice %>% filter(darq_pass == "FALSE") %>% ggplot(aes(x = dice_probability, y = darq_probability, color = darq_pass)) +
  geom_point()
```

## Explore
What if we only look at dice?

### histrogram on dice probability
Here is a histogram of dice probability, colored according to Pass/Fail based on DARQ.

```{r, echo = FALSE}
oasis_dice %>% ggplot(aes(dice_probability, fill = darq_pass)) +
geom_histogram(bins = 30)
```

### Regroup
Group the participants based on: <darq_pass>_<dice_pass>
```{r, echo = FALSE}
oasis_dice$dice_pass <- (oasis_dice$dice_probability > 0.91)

oasis_dice$both_pass <- sprintf("%s_%s", as.character(oasis_dice$darq_pass), as.character(oasis_dice$dice_pass))
```

### Clustering results
```{r, echo = FALSE}
oasis_dice_num <- oasis_dice[,c(4,5)]
oasis.cluster <- hclust(dist(scale(oasis_dice_num)), method = "ward.D2")
dend <- as.dendrogram(oasis.cluster)

# install.packages("dendextend")

color.ind <- c("TRUE_TRUE" = "#228B22", 
               "TRUE_FALSE" = "#20B2AA", 
               "FALSE_TRUE" = "#6959CD", 
               "FALSE_FALSE" = "#B22222")
color.ind.all <- recode(as.character(oasis_dice$both_pass), !!!color.ind)
labels_colors(dend) <- color.ind.all

plot(dend)
colored_bars(color.ind.all[oasis.cluster$order], dend, rowLabels = "Groups")
legend(x = 50, y = 15, names(color.ind), fill = color.ind, col = color.ind, cex = 1, title = "darq_dice")
```


### PCA
```{r, echo = FALSE}
## run PCA
PCAres <- epPCA(oasis_dice_num, 
                DESIGN = oasis_dice$darq_pass,
                scale = "SS1",
                graphs = FALSE)

map.label <- createxyLabels.gen(lambda = PCAres$ExPosition.Data$eigs,
                                tau = PCAres$ExPosition.Data$t)

## Plot factor scores
fi.map <- createFactorMap(PCAres$ExPosition.Data$fi,
                          col.points = color.ind.all,
                          alpha.points = 0.3,
                          col.background = NULL,
                          col.axes = "black")

fi.map$zeMap_background + fi.map$zeMap_dots + map.label

## Column factor scores
fj.map <- createFactorMap(PCAres$ExPosition.Data$fj,
                          col.background = NULL,
                          col.axes = "black")

fj.arrow <- addArrows(PCAres$ExPosition.Data$fj)

fj.map$zeMap + fj.arrow + map.label
```




